Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2011-12-18 extended_catalog/paper/extended_catalog.tex
update

The capability to spatially resolve \gev $\gamma$-ray
sources is important for several reasons.  
Finding a coherent source extension across different energy bands can
help to associate a LAT source to an otherwise confused counterpart.
Furthermore, $\gamma$-rays emission from dark matter
annihilation has been predicted to be detectable by the LAT.
Some of the dark matter substructure in our Galaxy 
could be spatially resolvable by the LAT \citep{pre_luanch_dark_matter_fermi}.  
Characterization of spatial extension could help to identify this substructure.
Also,
due to the strong energy dependence of the LAT PSF, the spatial and
spectral characterization of a source cannot be
decoupled. An inaccurate
spatial model will bias the spectral model of the source and vice versa. Specifically,
modeling a spatially extended source as point-like will systematically
soften measured spectra. Furthermore, correctly
modeling source extension is important for 
understanding an entire region of the sky. For example,
an imperfect model of the spatially extended LMC introduced
significant residuals in the surrounding region \citep{first_cat,second_cat}.
Such residuals can bias the significance and measured spectrum of
neighboring sources in the densely populated Galactic plane.

@@

The significance of the extension of a source can be calculated from the
likelihood ratio test of a hypothesis with a spatially extended source and
a hypothesis with a point-like source. The test statistic for this procedure
is defined as
\begin{equation}
  \tsext=2\log(\likelihood_\text{ext}/\likelihood_\text{ps}) 
\end{equation}
where \likelihood is the Poisson likelihood.
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point-like source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
Section~\ref{monte_carlo_validation}.

@@

There will always be a small numerical discrepancy between the expected
photon distribution derived from a true point-like source and a very small
extended source due to numerical error in the convolution.  In most
situations, this error is insignificant.  But in particular for
very bright sources, this numerical error has the potential to bias the
test statistic for the extension test. Therefore, when calculating
\tsext, we compare the likelihood fitting the source with an extended
spatial model to the likelihood when the extension is
fixed to a very small value ($10^{-10}$ degrees in radius 
for a uniform disk model).

@@

We evaluated the false detection probability for the measurement of source extent
by fitting the extensions of point-like sources.  \cite{mattox_egret}
discuss that the test statistic distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the test statistic in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

Finally, Figure~\ref{time_sensitivity} shows the projected
detection threshold of the LAT to extension with a 10 year
exposure against 10 times
the isotropic background measured by EGRET. This background is
representative of the background near the Galactic plane.  For small
extended sources, the threshold improves by a factor larger
than the square root of the exposure because the LAT is signal-limited
at high energies where we are most sensitive to extension. For large
extended sources, the relevant background is over a larger spatial range
and so the improvement is closer to a factor corresponding
to the square root of the exposure and is caused by Poisson fluctuations in the background.

@@

It is impossible to discriminate using only LAT data between a
spatially extended source and multiple point-like sources separated by
angular distances comparable to or smaller than the size of the LAT PSF.
To assess the plausibility of source confusion for situations
in which $\tsext\ge16$, we developed an algorithm in \pointlike to
simultaneously fit the position and spectrum of two point-like sources.

@@

As further validation of our method for studying
spatially extended sources, we reanalyzed the twelve spatially extended
sources which were included in the 2FGL catalog \citep{second_cat}.  
Even though these sources had all been the subjects of dedicated
analyses and separate publications, and had been fit with a variety
of spatial models,
it is valuable to show that
these sources are significantly extended using our systematic 
method assuming radially-symmetric uniform disk spatial models.  On the other hand, for some of
these sources a uniform disk spatial model does not well describe the
observed extended emission and so the dedicated LAT publications provides
a qualitatively better model of these sources.

@@

Six extended SNRs were
included in the 2FGL catalog: W51C, IC~443, W28, W30, W44, the Cygnus Loop,
and W30
\citep{w51c,ic443,w28,w44,cygnus_loop_lat,w30_lat}.
Using photons
with energies between
1 \gev and 100 \gev, our analysis significantly detected
the extension of all six of these SNRs.

@@

Two nearby satellite galaxies of the Milky Way the Large Magellanic Cloud (LMC)
and the Small Magellanic
Cloud (SMC) were included in the 2FGL catalog as spatially extended sources \citep{lmc,smc}.  
Their extension was significantly
detected using photons with energies between
1 \gev and 100 \gev. Our
fit extension is comparable to the published result, but we note that
the previous LAT Collaboration publication on LMC used a more complicated two 2D Gaussian surface
brightness profile when fitting it \citep{lmc}.

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137, were identified
in the 2FGL catalog \citep{msh1552,velax,fermi_hess_j1825}.  HESS\,J1825$-$137
was significantly detected using photons with energies between 10
\gev and 100 \gev.  To avoid confusion with the nearby bright pulsar
PSR\,J1509$-$5850, MSH\,15$-$52 had to be analyzed at high energies.
Using photons with energies above 10 \gev, we fit the extension of
MSH\,15$-$52 to be consistent with the published size but with \tsext=6.6.

@@

Our analysis of these sources is summarized in
Table~\ref{known_extended_sources}.  This table includes the best fit
position and extension of these sources when fitting them 
with a radially-symmetric uniform disk model. It also
includes the best fit spectral parameters for each source.  The position
and extension of Vela X and the Centaurus A Lobes were taken from
\cite{velax,cen_a_lat} and are included in this list for completeness.

@@

We developed two criteria for estimating systematic errors
on the extension of the sources.
First, we estimated a systematic error due to
uncertainty in our knowledge of the LAT PSF.  
Before launch, the LAT PSF
was determined by detector simulations which were verified in accelerator
beam tests \citep{atwood_LAT_mission}. However, in-flight data revealed
a discrepancy above a few \gev in the PSF compared to the angular
distribution of photons from bright AGN \citep{lat_on_orbit_psf}.
Subsequently, the PSF was fit empirically to bright AGN and this
empirical parameterization is used in the P7\_V6 IRFs.  To account for
the uncertainty in our knowledge of the PSF, we refit our extended source
candidates using the pre-flight Monte Carlo representation of the PSF
and consider the difference in extension found using the two PSFs as a
systematic error on the extension of a source.  The same approach was used
in \cite{ic443}.  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic error is conservative.

@@

We estimated a second systematic error on the extension of a source
due to uncertainty in our model of the Galactic diffuse emission by
using an alternative model which takes as input templates
calculated by
GALPROP\footnote{GALPROP is a software package for calculating the
Galactic $\gamma$-ray emission based on a model of cosmic-ray propagation
in the Galaxy \citep{galprop1998,galprop2011}. 
See also \url{http://galprop.stanford.edu/} for details.} 
but then fits each template locally in
the surrounding region.
The particular GALPROP model that was used as input is described in
the LAT analysis of the isotropic diffuse
emission \citep{isotropic_lat}.  
The intensities of various components
of the Galactic diffuse emission were fitted individually using a
spatial distribution predicted by
the model.  We distinguished contributions from cosmic-ray interactions with the
molecular hydrogen, the atomic and ionized hydrogen, residual gas traced
by dust \citep{isabelle_dark_gass}, and the interstellar radiation
field. We further split the contributions from interactions with molecular
and atomic hydrogen to the Galactic diffuse emission according to the
distance from the Galactic center in which they are produced. Hence, we
replaced the standard diffuse emission model by 18 individually fitted
templates to describe individual components of the diffuse emission.
A similar crosscheck was used in an analysis of RX\,J1713.7$-$3946 
by the LAT Collaboration \citep{rx_j1713_lat}.

It is not expected that this diffuse model is superior to the standard
LAT model obtained through an all-sky fit.  However, adding degrees of
freedom to the background model can remove likely spurious sources that
correlate with features in the Galactic diffuse emission.  Therefore,
this tests systematics that may be due to imperfect modeling of the
diffuse emission in the region. Nevertheless, this alternative diffuse
model does not test all systematics related to the diffuse emission model.
In particular, because the alternative model uses the same underlying gas
maps it is unable to assess systematics due to mischaracterization of local
diffuse structure due to lack of resolution of the underlying maps.

@@

But this is infeasible in the Galactic plane where we most likely expect
to discover new spatially extended sources.  In the Galactic plane,
this analysis is challenged by our imperfect model of the diffuse
emission and by an imperfect model of nearby sources.  The Monte Carlo
study in Section~\ref{dual_localization_method}
showed that the overall likelihood would greatly increase by fitting
a spatially extended source as two point-like sources so we expect
that spatially extended LAT sources are modeled in the 2FGL catalog as
multiple point-like sources. Furthermore, the position of other nearby sources
in the region close to an extended source could be biased by not correctly
modeling the extension of the source.  Since 2FGL produced a
list of sources significant at energies above 100 \mev whereas we are
most sensitive to spatial extension at higher energies,
we also expect that at higher energies our analysis would be complicated
by 2FGL sources no longer significant and by 2FGL
sources whose position was biased by diffuse emission at lower energies.

To account for these issues, we first produced a large list of possibly
extended sources employing very liberal search criteria and then
refined the analysis of the promising candidates on a case by case basis.
Our strategy tested all point-like 2FGL sources for extension assuming
they had a uniform radially-symmetric disk spatial model
and a power-law spectral model.  Although not all extended sources are
expected to have a shape very similar to a uniform disk, Section~\ref{compare_source_size} showed that for many spatially
extended sources the wide PSF of the LAT and limited statistics makes
this a reasonable approximation.  On the other hand, choosing this
spatial model biases us against finding extended sources that are not
well described by a uniform disk model such as shell-type supernova remnants.

@@

As was shown in Section~\ref{extension_sensitivity}, we gain little in
sensitivity using photons with energies below 1 \gev. On the other hand,
the broad PSF at low energy makes us more susceptible to systematic
errors arising from source confusion due to nearby soft point-like sources
and by uncertainties in our modeling of the Galactic diffuse emission. 
For that reason,
we performed our search using only photons with energies between 1 \gev
and 100 \gev.

We also performed a second search for extended sources using only
photons with energies between 10 \gev and 100 \gev.  Although this
approach tests the same sources, it is complementary because the Galactic
diffuse emission is even less dominant above 10 \gev and because source
confusion is less of an issue.  A similar procedure was used to detect
the spatial extension of MSH\,15$-$52 and
HESS\,J1825$-$137 with the LAT \citep{msh1552,fermi_hess_j1825}.

@@

Similarly, many of these sources were confused with nearby
point-like sources or influenced by large-scale residuals in the
diffuse emission.  To help determine which of these fits found
truly extended sourced and when the extension was
influenced by source confusion and diffuse emission, we generated a
series of diagnostic plots.  For each candidate, we generated a map
of the residual test statistic by adding a new point-like source of spectral index 2 into the
region at each position and finding the increase in likelihood when fitting
its flux. Figure~\ref{res_tsmaps} shows this map around the
most significantly extended source IC~443 when it is modeled both as a
point-like source and as an extended source.  The residual test statistic
map indicates
that the spatially extended model for IC~443 is a significantly better
description of the observed photons and that there is no significant
residual in the region after modeling the source as being spatially extended.
We also generated plots of the sum of all counts within a given distance of
the source and compared it to the model predictions assuming the emission
originated from a point-like source.  An example radial integral plot
is shown for the extended source IC~443 in Figure~\ref{four_plots_ic443}.
For each source, we also made diffuse-emission-subtracted smoothed counts
maps (shown for IC~443 in Figure~\ref{four_plots_ic443}).

We found by visual inspection that in many
cases our analysis was strongly influenced by large-scale residuals in the
diffuse emission and hence the extension measure was unreliable.  This was
especially true in our analysis of sources in the 1 \gev to 100 \gev
energy range.  An example of such a case is 2FGL\,J1856.2+0450c analyzed
in the 1 \gev to 100 \gev energy range. Figure~\ref{example_bad_fit}
shows a diffuse-emission-subtracted smoothed counts map for
this source with the best
fit extension of the source overlaid. There appears to be large-scale
residuals in the diffuse emission in this region along the Galactic plane.
As a result, 2FGL\,J1856.2+0450c is fit to an extension of $\sim2\degree$
and the result is statistically significant with \tsext=45.4. However,
by looking at the residuals it is clear that this complicated region is
not well modeled. We manually discard sources like this.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml
% (2FGL J1856.2+0450c = P72Y3047 for)

@@

For each candidate that was not biased by neighboring point-like
sources or by large-scale residuals in the diffuse emission model, we
improved the model of the region by deciding on a case by case basis which
background point-like sources should be kept.  We kept in our model of the sky
sources that we
believed represented physically distinct sources and we removed 
sources that we believed were included in the 2FGL catalog to compensate
for residuals induced by not modeling the extension of the source.
Soft nearby point-like 2FGL sources that were not significant at higher energies
were frozen to the spectrum predicted by 2FGL.
When deciding
which background sources to keep and which to remove, we used as much as possible
multiwavelength information about possibly extended source counterparts
to help guide our choice. For each extended source presented in 
Section~\ref{new_ext_srcs_section}, we describe any modification from 2FGL
of the background model that were performed.

The best fit position of nearby point-like sources can be influenced by
the extended source and vice versa.  Similarly,
the best fit position of nearby point-like sources in the 2FGL catalog can be biased
by systematic issues at lower energies.
Therefore, after selecting the list of background sources, we iteratively
refit the position and spectrum of nearby background sources as well as
the position and extension of the analyzed spatially extended
sources until the overall fit converged globally.  For each extended
source, we will describe the position of any relocalized background
sources.

After obtaining the overall best fit positions and extensions of all
of the sources in the region using \pointlike, we refit the spectral
parameters of the region using \gtlike.  With \gtlike, we obtained a
second measure of \tsext.  We only consider a source to be extended when
both \pointlike and \gtlike agree that $\tsext>16$. 
We further required that $\tsext>16$ using the alternative diffuse
model presented in Section~\ref{systematic_errors_on_extension}.
We then replaced the spatially extended source with two point-like
sources and refit the position and spectrum of the
two point-like sources to calculate \tsinc.
We only consider a source to be spatially extended, instead of being
the result of confusion of two point-like sources, if $\tsext>\tsinc$.
As was shown in Section~\ref{dual_localization_method}, this test is
fairly powerful at removing situations in which the emission actually
originates from two distinct point-like sources instead of one spatially
extended source.  On the other hand, it is still possible that improved
observations could resolve additional structure or new sources that we are
currently not sensitive to. Considering the very complicated morphologies
of extended sources observed at other wavelengths and the high density
of possible sources that are expected to emit at \gev energies, it is
likely that in some of these regions further observations will resolve
that the emission is significantly more complicated than the simple
radially-symmetric uniform disk model that we assume.

@@

Nine extended sources not included in the 2FGL catalog were found by our
extended source search. Two of these have been previously
studied in dedicated publications: RX\,J1713.7$-$3946 and Vela
Jr. \citep{rx_j1713_lat,vela_jr_lat}.
Two of these sources were
found when using photons with energies between 1 \gev and 100 \gev and seven
were found when using photons with energies between 10 \gev and 100 \gev.
For the sources found at energies above 10 \gev, we restrict our
analysis to higher energies because of the issues
of source confusion and diffuse emission modeling described in
Section~\ref{extended_source_search_method}.
The spectral and spatial properties of these nine sources are summarized
in Table~\ref{new_ext_srcs_table} and the results of our investigation of
systematic errors are presented in Table~\ref{alt_diff_model_results}.
Table~\ref{alt_diff_model_results}
also
compares the likelihood assuming the source is spatially
extended to the likelihood assuming 
that the emission originates from two independent point-like
sources. For these new extended sources, $\tsext>\tsinc$ 
so we conclude that 
the \gev emission does not originate from two physically
distinct point-like sources (see Section~\ref{dual_localization_method}).  
Table~\ref{alt_diff_model_results} also includes the
results of the extension fit using variations of the PSF and the Galactic
diffuse model described in Section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using the standard
analysis, the alternative diffuse models, and the alternative PSF.
This suggests that the sources are robust against features in the diffuse
emission model and uncertainties in the angular resolution of the LAT.

@@

This source is in a region of remarkably complicated diffuse emission.
Even though it is $16\degree$ from the Galactic plane, this source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI and significant 
dark gas found only by its association with dust emission
\citep{isabelle_dark_gass}. Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and 
${}^{12}\text{CO}$ ($J=1\rightarrow 0$)
contours are overlaid on Figure~\ref{1FGL_J1628.6-2419c} and show good
spatial correlation with the \gev emission \citep{iras_rho_ophiuci,co_rho_ophiuci}.
This source might 
represent $\gamma$-ray emission from the interactions of cosmic rays with
interstellar gas which has not been accounted for in the LAT diffuse
emission model.

@@

% Veritas observations of $\gamma$-Cygni: 
%   http://arxiv.org/pdf/0912.4492v1
%   extended: VER J019+407
%     Veritas observations of $\gamma$-Cygni veritas_gamma_cygni,

>>>>

The capability to spatially resolve \gev $\gamma$-ray
sources is important for several reasons.  
Finding a coherent source extension across different energy bands can
help to associate a LAT source to an otherwise confused counterpart.
Furthermore, $\gamma$-rays emission from dark matter
annihilation has been predicted to be detectable by the LAT.
Some of the dark matter substructure in our Galaxy 
could be spatially resolvable by the LAT \citep{pre_luanch_dark_matter_fermi}.  
Characterization of spatial extension could help to identify this substructure.
Also,
due to the strong energy dependence of the LAT PSF, the spatial and
spectral characterization of a source cannot be
decoupled. An inaccurate
spatial model will bias the spectral model of the source and vice versa. Specifically,
modeling a spatially extended source as point-like will systematically
soften measured spectra. Furthermore, correctly
modeling source extension is important for 
understanding an entire region of the sky. For example,
an imperfect model of the spatially extended LMC introduced
significant residuals in the surrounding region \citep{first_cat,second_cat}.
Such residuals can bias the significance and measured spectra of
neighboring sources in the densely populated Galactic plane.

@@

The significance of the extension of a source can be calculated from the
likelihood ratio test of a hypothesis with a spatially extended source and
a hypothesis with a point-like source. The test statistic (TS) for this procedure
is defined as
\begin{equation}
  \tsext=2\log(\likelihood_\text{ext}/\likelihood_\text{ps}) 
\end{equation}
where \likelihood is the Poisson likelihood.
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point-like source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
Section~\ref{monte_carlo_validation}.

@@

There will always be a small numerical discrepancy between the expected
photon distribution derived from a true point-like source and a very small
extended source due to numerical error in the convolution.  In most
situations, this error is insignificant.  But in particular for
very bright sources, this numerical error has the potential to bias the
TS for the extension test. Therefore, when calculating
\tsext, we compare the likelihood fitting the source with an extended
spatial model to the likelihood when the extension is
fixed to a very small value ($10^{-10}$ degrees in radius 
for a uniform disk model).

@@

We evaluated the false detection probability for the measurement of source extent
by fitting the extensions of point-like sources.  \cite{mattox_egret}
discuss that the TS distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the TS in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

Finally, Figure~\ref{time_sensitivity} shows the projected
detection threshold of the LAT to extension with a 10 year
exposure against 10 times
the isotropic background measured by EGRET. This background is
representative of the background near the Galactic plane.  For small
extended sources, the threshold improves by a factor larger
than the square root of the relative exposures because the LAT is signal-limited
at high energies where the present analysis is most sensitive. For large
extended sources, the relevant background is over a larger spatial range
and so the improvement is closer to a factor corresponding
to the square root of the relative exposures and is caused by Poisson fluctuations in the background.

@@

It is impossible to discriminate using only LAT data between a
spatially extended source and multiple point-like sources separated by
angular distances comparable to or smaller than the size of the LAT PSF.
To assess the plausibility of source confusion for situations
in which $\tsext\ge16$, we developed an algorithm in \pointlike to
simultaneously fit the positions and spectra of two point-like sources.

@@

As further validation of our method for studying
spatially extended sources, we reanalyzed the twelve spatially extended
sources which were included in the 2FGL catalog \citep{second_cat}.  
Even though these sources had all been the subjects of dedicated
analyses and separate publications, and had been fit with a variety
of spatial models,
it is valuable to show that
these sources are significantly extended using our systematic 
method assuming radially-symmetric uniform disk spatial models.  On the other hand, for some of
these sources a uniform disk spatial model does not well describe the
observed extended emission and so the dedicated LAT 
publications provide better models of these sources.

@@

Six extended SNRs were
included in the 2FGL catalog: W51C, IC~443, W28, W30, W44, the Cygnus Loop,
and W30
\citep{w51c,ic443,w28,w44,cygnus_loop_lat,w30_lat}.
Using photons
with energies between
1 \gev and 100 \gev, our analysis significantly detected
that these six SNRs are spatially extended.

@@

Two nearby satellite galaxies of the Milky Way the Large Magellanic Cloud (LMC)
and the Small Magellanic
Cloud (SMC) were included in the 2FGL catalog as spatially extended sources \citep{lmc,smc}.  
Their extensions were significantly
detected using photons with energies between
1 \gev and 100 \gev. Our
fit extensions are comparable to the published result, but we note that
the previous LAT Collaboration publication on the LMC used a more complicated two 2D Gaussian surface
brightness profile when fitting it \citep{lmc}.

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137, were fit as
extended sources in the 2FGL analysis \citep{msh1552,velax,fermi_hess_j1825}.  
In the present analysis, HESS\,J1825$-$137
was significantly detected using photons with energies between 10
\gev and 100 \gev.  To avoid confusion with the nearby bright pulsar
PSR\,J1509$-$5850, MSH\,15$-$52 had to be analyzed at high energies.
Using photons with energies above 10 \gev, we fit the extension of
MSH\,15$-$52 to be consistent with the published size but with \tsext=6.6.

@@

Our analysis of these sources is summarized in
Table~\ref{known_extended_sources}.  This table includes the best fit
positions and extensions of these sources when fitting them 
with a radially-symmetric uniform disk model. It also
includes the best fit spectral parameters for each source.  The positions
and extensions of Vela X and the Centaurus A Lobes were taken from
\cite{velax,cen_a_lat} and are included in this table for completeness.

@@

We developed two criteria for estimating systematic errors
on the extensions of the sources.
First, we estimated a systematic error due to
uncertainty in our knowledge of the LAT PSF.  
Before launch, the LAT PSF
was determined by detector simulations which were verified in accelerator
beam tests \citep{atwood_LAT_mission}. However, in-flight data revealed
a discrepancy above 3 \gev in the PSF compared to the angular
distribution of photons from bright AGN \citep{lat_on_orbit_psf}.
Subsequently, the PSF was fit empirically to bright AGN and this
empirical parameterization is used in the P7\_V6 IRFs.  To account for
the uncertainty in our knowledge of the PSF, we refit our extended source
candidates using the pre-flight Monte Carlo representation of the PSF
and consider the difference in extension found using the two PSFs as a
systematic error on the extension of a source.  The same approach was used
in \cite{ic443}.  we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic error is conservative.

@@

We estimated a second systematic error on the extension of a source
due to uncertainty in our model of the Galactic diffuse emission by
using an alternative 
approach to modeling the diffuse emission
which takes as input templates
calculated by
GALPROP\footnote{GALPROP is a software package for calculating the
Galactic $\gamma$-ray emission based on a model of cosmic-ray propagation
in the Galaxy and maps of the distributions of the 
components of the interstellar medium \citep{galprop1998,galprop2011}. 
See also \url{http://galprop.stanford.edu/} for details.} 
but then fits each template locally in
the surrounding region.
The particular GALPROP model that was used as input is described in
the LAT analysis of the isotropic diffuse
emission \citep{isotropic_lat}.  
The intensities of various components
of the Galactic diffuse emission were fitted individually using a
spatial distribution predicted by
the model.  We distinguished contributions from cosmic-ray interactions with the
molecular hydrogen, the atomic and ionized hydrogen, residual gas traced
by dust \citep{isabelle_dark_gass}, and the interstellar radiation
field. We further split the contributions from interactions with molecular
and atomic hydrogen to the Galactic diffuse emission according to the
distance from the Galactic center in which they are produced. Hence, we
replaced the standard diffuse emission model by 18 individually fitted
templates to describe individual components of the diffuse emission.
A similar crosscheck was used in an analysis of RX\,J1713.7$-$3946 
by the LAT Collaboration \citep{rx_j1713_lat}.

It is not expected that this diffuse model is superior to the standard
LAT model obtained through an all-sky fit.  However, adding degrees of
freedom to the background model can remove likely spurious sources that
correlate with features in the Galactic diffuse emission.  Therefore,
this tests systematics that may be due to imperfect modeling of the
diffuse emission in the region. Nevertheless, 
this alternative approach to modeling the diffuse emission
does not test all systematics related to the diffuse emission model.
In particular, because the alternative approach uses the same underlying gas
maps it is unable to be used to assess systematics due to mischaracterization of local
diffuse structure due to lack of resolution of the underlying maps.

@@

But this is infeasible in the Galactic plane where we most likely expect
to discover new spatially extended sources.  In the Galactic plane,
this analysis is challenged by our imperfect model of the diffuse
emission and by an imperfect model of nearby sources.  The Monte Carlo
study in Section~\ref{dual_localization_method}
showed that the overall likelihood would greatly increase by fitting
a spatially extended source as two point-like sources so we expect
that spatially extended sources would be modeled in the 2FGL catalog as
multiple point-like sources. Furthermore, the positions of other nearby sources
in the region close to an extended source could be biased by not correctly
modeling the extension of the source.  Since 2FGL produced a
list of sources significant at energies above 100 \mev whereas we are
most sensitive to spatial extension at higher energies,
we also expect that at higher energies our analysis would be complicated
by 2FGL sources no longer significant and by 2FGL
sources whose positions were biased by diffuse emission at lower energies.

To account for these issues, we first produced a large list of possibly
extended sources employing very liberal search criteria and then
refined the analysis of the promising candidates on a case by case basis.
Our strategy was to test all point-like 2FGL sources for extension assuming
they had a uniform radially-symmetric disk spatial model
and a power-law spectral model.  Although not all extended sources are
expected to have a shape very similar to a uniform disk, Section~\ref{compare_source_size} showed that for many spatially
extended sources the wide PSF of the LAT and limited statistics makes
this a reasonable approximation.  On the other hand, choosing this
spatial model biases us against finding extended sources that are not
well described by a uniform disk model such as shell-type SNRs.

@@

As was shown in Section~\ref{extension_sensitivity}, little sensitivity is gained by
using photons with energies below 1 \gev. On the other hand,
the broad PSF at low energy makes the analysis more susceptible to systematic
errors arising from source confusion due to nearby soft point-like sources
and by uncertainties in our modeling of the Galactic diffuse emission. 
For that reason,
we performed our search using only photons with energies between 1 \gev
and 100 \gev.

We also performed a second search for extended sources using only
photons with energies between 10 \gev and 100 \gev.  Although this
approach tests the same sources, it is complementary because the Galactic
diffuse emission is even less dominant above 10 \gev and because source
confusion is less of an issue.  A similar procedure was used to detect
the spatial extensions of MSH\,15$-$52 and
HESS\,J1825$-$137 with the LAT \citep{msh1552,fermi_hess_j1825}.

@@

Similarly, many of these sources were confused with nearby
point-like sources or influenced by large-scale residuals in the
diffuse emission.  To help determine which of these fits found
truly extended sources and when the extension was
influenced by source confusion and diffuse emission, we generated a
series of diagnostic plots.  For each candidate, we generated a map
of the residual TS by adding a new point-like source of spectral index 2 into the
region at each position and finding the increase in likelihood when fitting
its flux. Figure~\ref{res_tsmaps} shows this map around the
most significantly extended source IC~443 when it is modeled both as a
point-like source and as an extended source.  The residual TS
map indicates
that the spatially extended model for IC~443 is a significantly better
description of the observed photons and that there is no significant
residual in the region after modeling the source as being spatially extended.
We also generated plots of the sum of all counts within a given distance of
the source and compared them to the model predictions assuming the emission
originated from a point-like source.  An example radial integral plot
is shown for the extended source IC~443 in Figure~\ref{four_plots_ic443}.
For each source, we also made diffuse-emission-subtracted smoothed counts
maps (shown for IC~443 in Figure~\ref{four_plots_ic443}).

We found by visual inspection that in many
cases our results were strongly influenced by large-scale residuals in the
diffuse emission and hence the extension measure was unreliable.  This was
especially true in our analysis of sources in the 1 \gev to 100 \gev
energy range.  An example of such a case is 2FGL\,J1856.2+0450c analyzed
in the 1 \gev to 100 \gev energy range. Figure~\ref{example_bad_fit}
shows a diffuse-emission-subtracted smoothed counts map for
this source with the best
fit extension of the source overlaid. There appear to be large-scale
residuals in the diffuse emission in this region along the Galactic plane.
As a result, 2FGL\,J1856.2+0450c is fit to an extension of $\sim2\degree$
and the result is statistically significant with \tsext=45.4. However,
by looking at the residuals it is clear that this complicated region is
not well modeled. We manually discard sources like this.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml
% (2FGL J1856.2+0450c = P72Y3047 for)

@@

For each candidate that was not biased by neighboring point-like
sources or by large-scale residuals in the diffuse emission model, we
improved the model of the region by deciding on a case by case basis which
background point-like sources should be kept.  We kept in our model the
sources that we
believed represented physically distinct sources and we removed 
sources that we believed were included in the 2FGL catalog to compensate
for residuals induced by not modeling the extension of the source.
Soft nearby point-like 2FGL sources that were not significant at higher energies
were frozen to the spectras predicted by 2FGL.
When deciding
which background sources to keep and which to remove, we used as much as possible
multiwavelength information about possibly extended source counterparts
to help guide our choice. For each extended source presented in 
Section~\ref{new_ext_srcs_section}, we describe any modifications from 2FGL
of the background model that were performed.

The best fit positions of nearby point-like sources can be influenced by
the extended source and vice versa.  Similarly,
the best fit positions of nearby point-like sources in the 2FGL catalog can be biased
by systematic issues at lower energies.
Therefore, after selecting the list of background sources, we iteratively
refit the positions and spectra of nearby background sources as well as
the positions and extensions of the analyzed spatially extended
sources until the overall fit converged globally.  For each extended
source, we will describe the positions of any relocalized background
sources.

After obtaining the overall best fit positions and extensions of all
of the sources in the region using \pointlike, we refit the spectral
parameters of the region using \gtlike.  With \gtlike, we obtained a
second measure of \tsext.  We only consider a source to be extended when
both \pointlike and \gtlike agree that $\tsext>16$. 
We further required that $\tsext>16$ using the 
alternative approach to modeling the diffuse emission
presented in Section~\ref{systematic_errors_on_extension}.
We then replaced the spatially extended source with two point-like
sources and refit the positions and spectra of the
two point-like sources to calculate \tsinc.
We only consider a source to be spatially extended, instead of being
the result of confusion of two point-like sources, if $\tsext>\tsinc$.
As was shown in Section~\ref{dual_localization_method}, this test is
fairly powerful at removing situations in which the emission actually
originates from two distinct point-like sources instead of one spatially
extended source.  On the other hand, it is still possible that longer
observations could resolve additional structure or new sources 
that the analysis cannot currently detect. 
Considering the very complicated morphologies
of extended sources observed at other wavelengths and the high density
of possible sources that are expected to emit at \gev energies, it is
likely that in some of these regions further observations will reveal
that the emission is significantly more complicated than the simple
radially-symmetric uniform disk model that we assume.

@@

Nine extended sources not included in the 2FGL catalog were found by our
extended source search. Two of these have been previously
studied in dedicated publications: RX\,J1713.7$-$3946 and Vela
Jr. \citep{rx_j1713_lat,vela_jr_lat}.
Two of these sources were
found when using photons with energies between 1 \gev and 100 \gev and seven
were found when using photons with energies between 10 \gev and 100 \gev.
For the sources found at energies above 10 \gev, we restrict our
analysis to higher energies because of the issues
of source confusion and diffuse emission modeling described in
Section~\ref{extended_source_search_method}.
The spectral and spatial properties of these nine sources are summarized
in Table~\ref{new_ext_srcs_table} and the results of our investigation of
systematic errors are presented in Table~\ref{alt_diff_model_results}.
Table~\ref{alt_diff_model_results}
also
compares the likelihood assuming the source is spatially
extended to the likelihood assuming 
that the emission originates from two independent point-like
sources. For these new extended sources, $\tsext>\tsinc$ 
so we conclude that 
the \gev emission does not originate from two physically
distinct point-like sources (see Section~\ref{dual_localization_method}).  
Table~\ref{alt_diff_model_results} also includes the
results of the extension fits using variations of the PSF and the Galactic
diffuse model described in Section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using the standard
analysis, the alternative approach to modeling the diffuse emission, and the alternative PSF.
This suggests that the sources are robust against mis-modeled features in the diffuse
emission model and uncertainties in the PSF.

@@

@@

% VERITAS observations of $\gamma$-Cygni: 
%   http://arxiv.org/pdf/0912.4492v1
%   extended: VER J019+407
%     VERITAS observations of $\gamma$-Cygni veritas_gamma_cygni,
